### 2019-09-08 analysis/002_kb_rp_coverage.Rmd

---
title: "Reprint Author Analysis"
output: github_document
bibliography: pubs.bib
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)

options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r setup, echo = FALSE}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

## Questions

1. How many records in the Web of Science have affiliation information about reprint authors?

2. How many records in the Web of Science have more than one reprint author?

3. How many reprint authors are co-located (institutional level)?

4. How many reprint authors are internationally co-located (country level)?

5. What is the collaboration network in terms of country affiliations from reprint authors?

Focus:

- Database: wos_b_2019
- Document Types: Articles and Reviews
- Database Collections: `WOS.SCI`, `WOS.SSCI`, `WOS.AHCI`
- Publication Period 2014 - 2017

## Data analysis

### 1. How many records in the Web of Science have affiliation information at the country level about reprint authors?

#### Articles total

```{sql, connection=jdbcConnection, output.var="articles_total"}
select wos_b_2019.items.pubyear, COUNT(DISTINCT(wos_b_2019.items.ut_eid)) as articles_total
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017, 2018)
group by wos_b_2019.items.pubyear
```

```{r}
articles_total %>% 
  arrange(desc(PUBYEAR))
```

#### Records with reprint authors

Using `wos_b_2019.items_authors_institutions.type = 'RP'` to identify reprint authors.

```{sql, connection=jdbcConnection, output.var="rp_articles_total"}
select wos_b_2019.items.pubyear, COUNT(DISTINCT(wos_b_2019.items.ut_eid)) as rp_articles_total
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    inner join
        wos_b_2019.items_authors_institutions                                        
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items     
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017, 2018)              
        and wos_b_2019.items_authors_institutions.type = 'RP'
    group by wos_b_2019.items.pubyear
```

```{r}
rp_articles_total %>% 
  arrange(desc(PUBYEAR))
```

#### Reprint authors records with country information

*Retrieval considerations*

[An earlier exploration](kb_rp_exploration.md) studying how affiliations from reprint authors are represented in the WOS-KB revealed that only one affiliation is tagged with `RP in the table `wos_b_2019.items_author_institutions` per author. To obtain all affiliations per reprint author, the following retrieval strategy can be used:

1. Get all authors tagged as RP per item
2. Obtain all institutional addresses per rp_author-item combination including countrycode
3. Summarize observations

*Country codes*

There are two tables containing country codes, `wos_b_2019.institutions.countrycode` and `wos_b_2019.d_items_authors_institutions.inst_countrycode`. First, *`wos_b_2019.institutions.countrycode`* is analyzed

```{sql, connection=jdbcConnection, output.var="rp_cc_articles_total"}
select
        d.pubyear, count(distinct(d.ut_eid)) as rp_cc_articles_total
    from
        wos_b_2019.items_authors_institutions 
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear
            from
                wos_b_2019.items     
            inner join
                wos_b_2019.databasecollection                           
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.items_authors_institutions                                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'
                )                           
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'
                )                            
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018 
                )                       
                and wos_b_2019.items_authors_institutions.type = 'RP'         
        ) d 
            on (
                wos_b_2019.items_authors_institutions.fk_items = d.pk_items 
                and wos_b_2019.items_authors_institutions.fk_authors = d.fk_authors
            )
         inner join wos_b_2019.institutions on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions
         where wos_b_2019.institutions.countrycode is not null     
   group by d.pubyear
```

```{r}
rp_cc_articles_total %>% 
  arrange(desc(PUBYEAR))
```

Data Summary:

```{r}
rq_1_df <- articles_total %>%
  left_join(rp_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_RP_RATIO = RP_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  left_join(rp_cc_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_CC_RP_RATIO = RP_CC_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  arrange(desc(PUBYEAR))
rq_1_df
```

```{r}
rq_1_df %>%
  gather(ARTICLES_TOTAL, RP_ARTICLES_TOTAL, RP_CC_ARTICLES_TOTAL, 
         key = "indicator", value = "wos_articles") %>%
  ggplot(aes(PUBYEAR, wos_articles, fill = indicator)) +
  geom_bar(position="dodge", stat="identity") +
  scale_fill_viridis_d("", option = "C", labels = c("Articles + Reviews", "with RP", "with RP and countrycode")) +
  theme_minimal() +
  labs(y = "wos_b_2019 records", 
       title = "Coverage of reprint authors with country of affiliation",
       subtitle = "wos_b_2019.institutions.countrycode") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Next, *`wos_b_2019.d_items_authors_institutions.inst_countrycode`* is analyzed.

```{sql connection=jdbcConnection, output.var="rp_d_cc_articles_total"}
  select
        d.pubyear, count(distinct(d.ut_eid)) as rp_d_cc_articles
    from
        wos_b_2019.d_items_authors_institutions 
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear
            from
                wos_b_2019.items     
            inner join
                wos_b_2019.databasecollection                           
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items             
            inner join
                wos_b_2019.items_authors_institutions                                                     
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'
                )                           
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'
                )                            
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018
                )                       
                and wos_b_2019.items_authors_institutions.type = 'RP'         
        ) d 
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items 
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors
            )
         where wos_b_2019.d_items_authors_institutions.inst_countrycode is not null     
   group by d.pubyear
```

```{r}
rp_d_cc_articles_total
```

```{r}
rp_df_d <- articles_total %>%
  left_join(rp_articles_total, by = "PUBYEAR") %>%
  mutate(ARTICLE_RP_RATIO = RP_ARTICLES_TOTAL / ARTICLES_TOTAL) %>%
  left_join(rp_d_cc_articles_total, by = "PUBYEAR") %>%
  mutate(RP_D_CC_ARTICLES_RATIO = RP_D_CC_ARTICLES / ARTICLES_TOTAL) %>%
  arrange(desc(PUBYEAR))
rp_df_d
```

```{r}
rp_df_d %>%
  gather(ARTICLES_TOTAL, RP_ARTICLES_TOTAL, RP_D_CC_ARTICLES, 
         key = "indicator", value = "wos_articles") %>%
  ggplot(aes(PUBYEAR, wos_articles, fill = indicator)) +
  geom_bar(position="dodge", stat="identity") +
  scale_fill_viridis_d("", option = "C", labels = c("Articles + Reviews", "with RP", "with RP and countrycode")) +
  theme_minimal() +
  labs(y = "wos_b_2019 records", 
       title = "Coverage of reprint authors with country of affiliation",
       subtitle = "wos_b_2019.d_items_authors_institutions.inst_countrycode") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

In conclusion, joining with both tables seems feasible, because it results in a thorough countries of affiliation coverage for reprint authors. Overall, `r round(mean(rp_df_d$RP_D_CC_ARTICLES_RATIO) * 100, 2)` % of records representing original articles and reviews in the period 2014 - 2017 provide affiliation information for reprint authors at the country-level. In the following, `wos_b_2019.d_items_authors_institutions.inst_countrycode` is used.

### 2. How many records in the Web of Science have more than one reprint author?

#### Query

```{sql connection=jdbcConnection, output.var="rp_ind_df"}
select
        distinct         
        d.pubyear,
        d.ut_eid,
        wos_b_2019.d_items_authors_institutions.fk_authors,
        wos_b_2019.d_items_authors_institutions.inst_per_auth_cnt,
        wos_b_2019.d_items_authors_institutions.country_per_auth_cnt     
    from
        wos_b_2019.d_items_authors_institutions      
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear             
            from
                wos_b_2019.items                  
            inner join
                wos_b_2019.databasecollection                                                
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                          
            inner join
                wos_b_2019.items_authors_institutions                                                                          
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items                               
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                 
                )                                            
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                 
                )                                             
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                 
                )                                        
                and wos_b_2019.items_authors_institutions.type = 'RP'                  
        ) d              
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                  
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors             
            )          
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null
```

#### Backup in KB table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_ind_df", 
             value = rp_ind_df,
             overwrite = TRUE)
```

#### Analysis: Distribution of the number of reprint authors

```{sql connection=jdbcConnection, output.var="rp_au"}
select
        pubyear,
        ut_eid,
        count(fk_authors) as n_au
    from
        rp_ind_df 
    group by
        ut_eid,
        pubyear 
```

Summary statistics

```{r}
summary(rp_au$N_AU)
```

Frequency table of the number of reprint authors per publication. Publications with more than five reprint authors per publication were reduced to the residual category "> 5".

```{r}
tp_au_count <- rp_au %>% 
  mutate(N_AU = fct_other(factor(N_AU), keep = c("1", "2", "3", "4", "5"),
                          other_level = "> 5")) %>% 
  group_by(N_AU, PUBYEAR) %>% 
  summarize(N = n()) %>% 
  ungroup() %>% 
  group_by(PUBYEAR) %>% 
  mutate(PROP = N / sum(N))
tp_au_count %>%
  ggplot(aes(N_AU, PROP, fill = factor(PUBYEAR))) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ factor(PUBYEAR)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of reprint authors per record",
    y = "Proportion",
    title = "Number of reprint authors per Web of Science record",
    subtitle = "Articles and reviews 2014 - 2018"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

SQL statement to retrieve all publications with more than one reprint author (incl. country info)

```{sql connection=jdbcConnection, output.var="rp_au_count"}
select
        pubyear,
        ut_eid,
        count(fk_authors)  
    from
        rp_ind_df 
    group by
        ut_eid,
        pubyear 
    having
        count(fk_authors) > 1
```

In total, `r length(unique(rp_au_count$UT_EID))` out of `r length(unique(rp_ind_df$UT_EID))` records indexed in the Web of Science for the period 2014 - 2017 have more than one reprint author, representing a percentage of `r round(length(unique(rp_au_count$UT_EID)) / length(unique(rp_ind_df$UT_EID)) * 100, 2)` %. Furthermore, data exploration suggests that starting from 2016 the Web of Science has begun to systematically keep track of more than one reprint author.

### 3. How many reprint authors are co-located (institutional level)?

#### Query

```{sql connection=jdbcConnection, output.var="rp_inst_au_count"}
select
        inst_per_auth_cnt,
        pubyear,
        count(*) as inst_n
    from
        rp_ind_df 
    group by
        pubyear,
        inst_per_auth_cnt
```

#### Analysis

Distribution of the number of institutional affiliations per reprint author and year

```{r}
rp_inst_au_df <- rp_inst_au_count %>%
  mutate(INST_PER_AUTH_CNT = fct_other(
    factor(INST_PER_AUTH_CNT),
    keep = c("1", "2", "3", "4", "5"),
    other_level = "> 5"
  )) %>%
  ungroup() %>%
  group_by(PUBYEAR) %>%
  mutate(PROP = INST_N / sum(INST_N)) %>%
  arrange(INST_PER_AUTH_CNT, PUBYEAR)

rp_inst_au_df %>%
  ggplot(aes(
    x = factor(INST_PER_AUTH_CNT),
    PROP,
    fill = factor(PUBYEAR)
  )) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ PUBYEAR) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of Institutions",
    y = "Proportion",
    title = "Institutional affiliations per reprint author",
    caption = "Derived from wos_b_2019.D_ITEMS_AUTHORS_INSTITUTIONS.INST_PER_AUTH_CNT"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Overall, around half of the reprint authors indexed listed more than one institutional affiliation (N = `r rp_inst_au_count %>% filter(!INST_PER_AUTH_CNT == "1") %>% .$INST_N %>% sum()`, Proportion = `r round(rp_inst_au_count %>% filter(!INST_PER_AUTH_CNT == "1") %>% .$INST_N %>% sum() / rp_inst_au_count  %>% .$INST_N %>% sum(), 4) * 100` %).

### 4. How many reprint authors are internationally co-located (country level)?

#### Query

```{sql connection=jdbcConnection, output.var="rp_country_au_count"}
select
        country_per_auth_cnt,
        pubyear,
        count(*) as country_n
    from
        rp_ind_df 
    group by
        pubyear,
        country_per_auth_cnt
```

Frequency table of the number of country affiliation per reprint author and publication. Reprint authors with more than three countries of affiliation at the time of publication were reduced to the residual category "Other".

```{r}
rp_country_au_df <- rp_country_au_count %>%
  mutate(COUNTRY_PER_AUTH_CNT = fct_other(
    factor(COUNTRY_PER_AUTH_CNT),
    keep = c("1", "2", "3"),
    other_level = "> 3"
  )) %>%
  ungroup() %>%
  group_by(PUBYEAR) %>%
  mutate(PROP = COUNTRY_N / sum(COUNTRY_N)) %>%
  arrange(COUNTRY_PER_AUTH_CNT, PUBYEAR)
```

```{r}
rp_country_au_df
```

```{r}
rp_country_au_df %>%
  ggplot(aes(
    x = factor(COUNTRY_PER_AUTH_CNT),
    PROP,
    fill = factor(PUBYEAR)
  )) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  facet_grid( ~ PUBYEAR) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 5L)) +
  labs(
    x = "Number of Countries",
    y = "Proportion",
    title = "Country affiliations per reprint author",
    caption = "Derived from wos_b_2019.D_ITEMS_AUTHORS_INSTITUTIONS.COUNTRY_PER_AUTH_CNT"
  ) +
  theme_minimal() +
  theme(legend.position = "none") +
  theme(plot.margin = margin(30, 30, 30, 30)) +
  theme(panel.grid.minor = element_blank()) +
  theme(axis.ticks = element_blank()) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank())
```

Although the number of co-located reprint authors is growing, only a small proportion of reprint author is internationally co-located (N = `r rp_country_au_df %>% filter(!COUNTRY_PER_AUTH_CNT == "1") %>% .$COUNTRY_N  %>% sum()`, Proportion = `r round(rp_country_au_df %>% filter(!COUNTRY_PER_AUTH_CNT == "1") %>% .$COUNTRY_N  %>% sum() / rp_country_au_df  %>% .$COUNTRY_N  %>% sum(), 4) * 100` %).

### 5. What is the collaboration network in terms of country affiliations from reprint authors?

#### Query

Obtain publications with country affiliations

```{sql connection=jdbcConnection, output.var="rp_country_mat"}
select  distinct wos_b_2019.d_items_authors_institutions.INST_COUNTRYCODE,       
        d.ut_eid
    from
        wos_b_2019.d_items_authors_institutions      
    inner join
        (
            select
                pk_items,
                fk_authors,
                ut_eid,
                pubyear             
            from
                wos_b_2019.items                  
            inner join
                wos_b_2019.databasecollection                                                
                    on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                          
            inner join
                wos_b_2019.items_authors_institutions                                                                          
                    on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items                               
            where
                wos_b_2019.databasecollection.edition_value in (
                    'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                 
                )                                            
                and wos_b_2019.items.doctype in (
                    'Article', 'Review'                 
                )                                             
                and wos_b_2019.items.pubyear in (
                    2014, 2015, 2016, 2017, 2018                 
                )                                        
                and wos_b_2019.items_authors_institutions.type = 'RP'                  
        ) d              
            on (
                wos_b_2019.d_items_authors_institutions.fk_items = d.pk_items                  
                and wos_b_2019.d_items_authors_institutions.fk_authors = d.fk_authors             
            )          
    where
        wos_b_2019.d_items_authors_institutions.inst_countrycode is not null
```

#### Analysis

Publications with internationally co-located reprint author(s)

```{r}
co_pubs <- rp_country_mat %>% 
  group_by(UT_EID) %>%
  filter(n() > 1)
```

Country share

```{r}
all_count <- rp_country_mat %>%
  count(INST_COUNTRYCODE, sort = TRUE)
co_count <- co_pubs %>%
  ungroup() %>%
  count(INST_COUNTRYCODE, sort = TRUE, name = "n_co")
left_join(all_count, co_count, by = "INST_COUNTRYCODE") %>%
  mutate(prop = n_co / n * 100) 
```

Create bi-partite matrix

```{r}
#co_pubs_mat <- as.matrix(table(co_pubs$UT_EID, co_pubs$INST_COUNTRYCODE))
```

calculate unipartite matrix (country level)

```{r}
#co_pubs_mat_t <- t(co_pubs_mat) %*% co_pubs_mat
```

visualise network (tbc)

```{r}
# library(sna)
# sna::gplot(co_pubs_mat_t, 
#            gmode = "graph", 
#            vertex.cex = log(diag(co_pubs_mat_t)^0.3),
#            usearrows = FALSE)
```