### 2019-09-08 analysis/004_kb_fetch_publisher.Rmd

---
title: "Publisher League Global"
author: "Najko Jahn"
date: "9/5/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

## About

This executable report describes how data about the global market shares of publishers were retrieved from the Web of Science (WoS) in-house database hosted by the [German Competence Center for Bibliometrics (KB)](http://www.bibliometrie.info/). This document is written in R Markdown. If you have access to WoS-KB data infrastructure, you will be able to replicate the methods used by calling

```r
require(rmarkdown)
require(tidyverse)
require(RJDBC)
require(rJava)
require(writexl)
rmarkdown::render("kb_fetch_publisher_global.Rmd")
```

## Workflow

### Obtain publication data from the Web of Science database hosted by the German Competence Centre for Bibliometrics

#### Connect to database

```{r}
# deal with rJava memory allocation 
# https://stackoverflow.com/questions/34624002/r-error-java-lang-outofmemoryerror-java-heap-space
options(java.parameters = "-Xmx1024m")
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "~/ojdbc8.jar") # you may need to change the path to your db driver
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    # login credentials are stored in the .Rprofile file 
    Sys.getenv("kb_user"), 
    Sys.getenv("kb_pwd")
  ) 
```

#### Query

```{sql connection=jdbcConnection, output.var="rp_publisher"}
select
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn,
        count(distinct(ut_eid)) as pubs                         
    from
        wos_b_2019.items                                            
    inner join
        wos_b_2019.databasecollection                                                                                          
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items                                  
    inner join
        wos_b_2019.issues                                                        
            on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues                             
    where
        wos_b_2019.databasecollection.edition_value in (
            'WOS.SCI', 'WOS.SSCI', 'WOS.AHCI'                                                   
        )                                                                              
        and wos_b_2019.items.doctype in (
            'Article', 'Review'                                                   
        )                                                                               
        and wos_b_2019.items.pubyear in (
            2014, 2015, 2016, 2017, 2018                                                   
        )                                                  
    group by
        wos_b_2019.items.pubyear,
        wos_b_2019.issues.issn
```

```{r}
head(rp_publisher) %>%
  knitr::kable()
# dump
write_csv(rp_publisher, "data/publisher_league_14_18.csv")
```

### Get most frequent publisher and journal names from Crossref

Publisher names used in the Web of Science are ambigue, and sometimes journals can change the publisher. To obtain most frequent publisher and journals names, Crossref was queried for every distinct journal by ISSN as described in [`fetch_data.R`](fetch_data.R).

As a result, a [matching table between Crossref and the Web of Science journals](data/cr_wos.csv) was created. In the following, this table is joined with the WoS data.

```{r}
cr_journals <- readr::read_csv("data/cr_wos.csv") %>%
  # merge springer nature brands
  mutate(publisher = ifelse(grepl("Springer", publisher, fixed = FALSE), "Springer Nature", publisher))
rp_publisher <- readr::read_csv("data/publisher_league_14_18.csv")
issn_l <- readr::read_tsv("data/20190818.ISSN-to-ISSN-L.txt") %>%
  # manual fix Journal - American Water Works 
  add_row(ISSN = "2164-4535", `ISSN-L` = "0003-150X")
rp_df <- rp_publisher %>%
  left_join(issn_l, by = "ISSN") %>%
  left_join(cr_journals, by = c(`ISSN-L`= "issn_l")) %>%
  select(-issn) %>%
  distinct()
```

Crossref indexed `r rp_df %>% filter(!is.na(publisher)) %>% distinct(ISSN) %>% nrow() / rp_df %>% distinct(ISSN) %>% nrow() * 100` % of Web of Science journals.

The following table shows the number of journals per publisher.

```{r}
rp_df %>% 
  distinct(ISSN, publisher) %>% 
  count(publisher, sort = TRUE) %>% 
  mutate(prop = n /sum(n) * 100)
```

Next, the number and proportion of original articles and reviews published per publisher is presented:

```{r}
rp_df %>% 
  group_by(publisher) %>%
  summarise(n = sum(PUBS)) %>% 
  mutate(prop = n /sum(n) * 100) %>%
  arrange(desc(prop))
```

### 3. Obtain open access status information

Bruns et al. provide a matching table for fully open access journals, joining multiple sources into one dataset.

Bruns, A., Lenke, C., Schmidt, C., & Taubert, N. C. (2019). ISSN-Matching of Gold OA Journals (ISSN-GOLD-OA) 3.0. Bielefeld University. doi:[10.4119/unibi/2934907](https://doi.org/10.4119/unibi/2934907)

The linking ISSN is used for matching.

```{r}
u <- "https://pub.uni-bielefeld.de/download/2934907/2934908/ISSN_Gold-OA_3.0.csv"
bie_oa <- readr::read_csv(u) %>%
  # remove missing entries with missing ISSN_L
  filter(!is.na(ISSN_L))
# add info to rp_df
rp_df_oa <- rp_df %>%
  mutate(oa_journal = `ISSN-L` %in% bie_oa$`ISSN_L`)
```

The following table shows the number and proportion of open access journals

```{r}
rp_df_oa %>% 
  distinct(`ISSN-L`, oa_journal) %>% 
  count(oa_journal) %>%
  mutate(prop = n / sum(n))
```

### 4. Data coding and dump

`CSV`-based dump

```{r}
rp_df_oa %>%
  select(
    issn_wos = ISSN,
    publication_year = PUBYEAR,
    articles = PUBS,
    journal_title, publisher, oa_journal,
    issn_l = `ISSN-L`
    ) %>%
  write_csv("data/publisher_14_18.csv")
```

Dump to Excel

```{r}
readr::read_csv("data/publisher_14_18.csv") %>%
  writexl::write_xlsx("data/publisher_14_18.xlsx")
```

Data Schema:

Data Schema:

|Variable           |Description                                                                                                                                                                      |Source                                                                                                                                                       |
|:------------------|:--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:------------------------------------------------------------------------------------------------------------------------------------------------------------|
|`issn_wos`         |ISSN, a standardized journal id.                                                                                                                                                 |KB Web of Science: `wos_b_2019.issues.issn`                                                                                                                  |
|`publication_year` |Year of publication, obtained from KB Web of Science                                                                                                                             |KB Web of Science: `wos_b_2019.items.pubyear`                                                                                                                |
|`articles`         |Number of original articles and reviews published. |KB Web of Science: Grouped counts over `wos_b_2019.issues.issn` and  `wos_b_2019.items.pubyear` |
|`journal_title`    |Most frequently used journal title in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                       |Crossref                                                                                                                                                     |
|`publisher`        |Most frequently used publisher name in terms of articles published between 2014 - 2018. If missing, the journal was not indexed in Crossref                                      |Crossref                                                                                                                                                     |
|`oa_journal`       |Is the journal publishing all articles open access without delay (full open access)?                                                                                             |Bielefeld GOLD OA List V3                                                                                                                                    |
|`issn_l`           |Linking ISSN, a journal id that groups the different media of the same serial publication, e.g. ISSN for print with electronic issn.                                             |CIEPS                                                                                                                                                        |