### 2019-10-22 analysis/001_kb_rp_pretest.Rmd

---
title: "Reprint Authors per Country"
output: github_document
bibliography: pubs.bib
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE
)
```

```{r setup, echo = FALSE}
require(tidyverse)
require(RJDBC)
require(rJava)
.jinit()
jdbcDriver <-
  JDBC(driverClass = "oracle.jdbc.OracleDriver", classPath = "../inst/jdbc_driver/ojdbc8.jar")
jdbcConnection <-
  dbConnect(
    jdbcDriver,
    "jdbc:oracle:thin:@//biblio-p-db01:1521/bibliodb01.fiz.karlsruhe",
    Sys.getenv("kb_user"),
    Sys.getenv("kb_pwd")
  ) 
```

## Motivation

Country information about corresponding authors play a crucial role in open access funding [@Schimmer_2015]. The Web of Science contains data about this author role including affiliations, making it a valuable source to determine the productivity of corresponding authors per country and journal.

The aim of this work is to better understand how countries of affiliation from corresponding authors are represented in the WoS-KB, and how they can be retrieved from it. Such an analysis will yield valuable insights about how to calculate the number of scholarly articles per country and journal from the WoS-KB database using the corresponding author addresses only.

## Potential issues

1. A corresponding author lists at least two different countries
2. More than one corresponding author is listed
3. Corresponding authors work in different countries

<!--  why is issue of ex. 1 (multiple affiliations with differing author roles) not mentioned here?

Maybe add something like

4. Multiple country affiliations of the same author are listed with differing author roles

The first three issues are of a more conceptual nature, all leading to a situation where one article is associated with multiple countries of affiliation. One has to decide upon a way of handling them which has to be considered at the time of interpretation of the results. Two options for the mentioned issues are whole counting or fractional counting. For whole counting, each article is counted fully for any country of affiliation of any listed corresponding author, so that the number of articles in the single categories do not necessarily add up to the total number of articles any more. Fractional counting assigns a fraction of one article count to each corresponding author, leading to possible non-integer total article counts. The second issue is not much of a problem in our setting, if both authors are situated in the same country, since we are interested in the country of affiliation associated with an article. If, however, the corresponding authors' countries of affiliation differ as in issue 3, we again have to decide upon a counting scheme: do we count each country only once per article or multiple times if multiple corresponding authors work there? The fourth issue is rather a problem of data consistency. -->

## Example 1

Potential issues: A corresponding author lists at least two different countries, e.g., `UT=000400754000138`. Additionally, only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.

```{sql connection=jdbcConnection}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000400754000138'
order by wos_b_2019.items_authors_institutions.fk_authors
```

## Example 2

Potential issues: More than one corresponding author is listed. For `UT_EID=000372645900002`, the Web of Science lists seven corresponding authors and four reprint adresses. Furthermore, as in Example 1, only one affiliation is tagged as `RP` in `wos_b_2019.items_author_institutions`.

```{sql connection=jdbcConnection, max.print = NA}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000372645900002'
order by wos_b_2019.items_authors_institutions.fk_authors
```

## Example 3

Corresponding authors work in different countries. For `UT_EID=000372645900002`, the Web of Science lists two corresponding authors, one affilated with a Spanish, the other with a Portuguese institution. As in Examples 1 and 2, only one affiliation is tagged with RP in `wos_b_2019.items_author_institutions`. Moreover, checking the WoS raw data reveals that the RP-author `14734495` is not affiliated with an Portuguese institution at all. <!--when I run the query it only shows 10 records, none of them has fk_author 14734495-->

```{sql connection=jdbcConnection}
select  distinct wos_b_2019.institutions.countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.items.ut_eid,
        wos_b_2019.items.pubyear
from wos_b_2019.items
inner join
        wos_b_2019.items_authors_institutions                                 
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions  
where wos_b_2019.items.ut_eid = '000389110200022'
order by wos_b_2019.items_authors_institutions.fk_authors
```

## SQL strategy to circumvent these issues

### Steps

1. Get all `FK_AUTHORS` tagged as `RP`, and store the result as `rp_author_table`
2. From authors listed in `rp_author_table` obtain all countries of affiliation and store them in `rp_countries_table`
3. Aggregate `rp_countries_table`

#### 1. Get all `FK_AUTHORS` tagged as `RP`, and store the result as `rp_author_table`

```{sql connection=jdbcConnection, output.var="rp_authors"}
select
        distinct wos_b_2019.items_authors_institutions.fk_items,
        wos_b_2019.items_authors_institutions.fk_authors      
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.databasecollection              
            on wos_b_2019.databasecollection.fk_items =  wos_b_2019.items.pk_items        
    inner join
        wos_b_2019.items_authors_institutions                                   
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
    inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions      
    where
        wos_b_2019.databasecollection.edition_value in ('WOS.SCI', 'WOS.SSCI', 'WOS.AHCI')                  
        and wos_b_2019.items.doctype in ('Article', 'Review')                   
        and wos_b_2019.items.pubyear in (2014, 2015, 2016, 2017)         
        and wos_b_2019.items.ut_eid in ('000389110200022', '000372645900002', '000400754000138') 
        and wos_b_2019.items_authors_institutions.type = 'RP'
```

Store in personal table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_author_table", 
             value = rp_authors)
```

#### From authors listed in `rp_author_table` obtain all countries of affilation and store them in `rp_countries_table`

```{sql connection=jdbcConnection, output.var="rp_countries"}
select
        countrycode,
        wos_b_2019.items_authors_institutions.type,
        wos_b_2019.items_authors_institutions.fk_institutions,
        wos_b_2019.items_authors_institutions.fk_authors,
        wos_b_2019.issues.issn,
        ut_eid,
        wos_b_2019.items.pubyear
    from
        wos_b_2019.items      
    inner join
        wos_b_2019.issues                           
            on wos_b_2019.issues.pk_issues = wos_b_2019.items.fk_issues         
    inner join
        wos_b_2019.items_authors_institutions                                        
            on wos_b_2019.items_authors_institutions.fk_items = wos_b_2019.items.pk_items       
    inner join
        wos_b_2019.institutions                          
            on wos_b_2019.institutions.pk_institutions = wos_b_2019.items_authors_institutions.fk_institutions      
    inner join
        rp_author_table                          
            on (
                rp_author_table.fk_items = wos_b_2019.items_authors_institutions.fk_items                  
                and rp_author_table.fk_authors = wos_b_2019.items_authors_institutions.fk_authors             
            )
```

Store in personal table space

```{r}
dbWriteTable(conn = jdbcConnection, 
             name = "rp_countries_table", 
             value = rp_countries)
```

#### Aggregate `rp_countries_table`

Counting method: country of affilation is only counted once per article.

```{sql connection=jdbcConnection}
select
    countrycode,
    issn,
    count(distinct(ut_eid)) as publications 
from
    rp_countries_table 
group by
    countrycode,
    issn
order by issn
```